---
output: github_document
---

<!--  README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  fig.width=6, 
  fig.height=4
)
```

# HeckmanStan

<!--[![CRAN_Status_Badge]()]()-->


The Heckman selection model is widely employed in econometric analysis and other social sciences, to address sample selection issues in data modeling. Its fundamental assumption hinges on the normality of error terms in both the outcome and the sample selection equations.

In this work, we explore these selection models using a Bayesian approach facilitated by the Stan software. Our objective is to compare the performance of the normal, Student’s-t and contaminated normal, aiming to address the complexities introduced by non-standard error distributions. 

<!--For more information about the model formulation and estimation, please see
Schumacher, F. L., Lachos, V. H., and Matos, L. A. (2021). Scale mixture of skew‐normal linear mixed models with within‐subject serial dependence. *Statistics in Medicine*. DOI: [10.1002/sim.8870](https://doi.org/10.1002/sim.8870).-->

## Installation

<!-- You can install the released version of lmmsmsn from [CRAN](https://CRAN.R-project.org) with: -->


You can install HeckmanStan from GitHub with:

```r
remotes::install_github("heeju-lim/HeckmanStan")
```
Or you can install the released version of skewlmm from [CRAN](https://CRAN.R-project.org) with:
```r
install.packages("HeckmanStan")
```

## Example

This is a basic example which shows you how to fit the Heckman Stan Model when data is generated from the normal distribution and fit the Heckmans Stan model using the contaminated normal distribution. 

```{r example1, out.width = '70%', fig.align = "center"}
library(HeckmanStan)
n<- 100
w<- cbind(1,rnorm(n),rnorm(n))
x<- cbind(w[,1:2])
type="CN"
sigma2<- 1
rho<-0.7
beta<- c(1,0.5)
gama<- c(1,0.3,-.5)
nu=c(0.1,0.1)
data<-geraHeckman(x,w,beta,gama,sigma2,rho,nu,type=type)
y<-data$y
cc<-data$cc
# Fit Heckman Normal Stan model
fit.n_stan <- HeckmanStan(y, x, w, cc, family="CN", thin = 5, chains = 1, iter = 1000, warmup = 100)
qoi=c("beta","gamma","sigma_e","sigma2", "rho","EAIC","EBIC")
print(fit.n_stan[[1]],par=qoi)
print(fit.n_stan[[2]])

# Plots for stanfit objects : 
library(rstan)
plot(fit.n_stan[[1]], pars=c("beta[1]","beta[2]", "gamma[1]", "gamma[2]", "gamma[3]", "rho", "sigma_e"))
plot(fit.n_stan[[1]], plotfun="hist", pars=c("beta[1]","beta[2]", "gamma[1]", "gamma[2]", "gamma[3]", "rho", "sigma_e"))
plot(fit.n_stan[[1]], plotfun="trace", pars=c("beta[1]","beta[2]", "gamma[1]", "gamma[2]", "gamma[3]", "rho", "sigma_e"))
```


